---
title: 'Lab #1'
author: "WEI-XUAN HU"
date: "2024-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## create a named numerical variable and a new vector of numbers.

```{r}
a <- 5
b <- 17
d <- c(1, 2, 3, 4, 5)
```

## Report these produced data objects and then apply at least two functions to each.
Here would be some stuff to describe them.

```{r}
sum(a, b)
mean(a, b)

sd(d)
max(d)
length(d)
```

## Create a new data frame with three columns and rows.

```{r tri_df_data, results='asis'}
library(knitr)
tri_df <- data.frame(name = c("SpongeBob", "Tom&Jerry", "Sesame Street"),
                     character = c("Sandy", "Jerry", "Cookie Monster"),
                     color = c("brown", "brown", "blue"),
                     finger_num = c(5, 5, 2))
kable(tri_df)
```
                     
## Load a dataset from the R package and report summary information on the dataset and its structure.

```{r}
library(MASS)
head(ChickWeight)
```
                
     
## Use the ls() function to see that the objects are stored in your workspace. Clear your workspace using the rm(list = ls()) function. Then, run the ls() function to make sure the objects are gone.     


## Sign up for a free student account on github and include your github username in your Markdown file.


                     
                     

