---
title: 'Lab #3'
author: "WEI-XUAN HU"
date: "2024-02-05"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## 0. Download packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# There are some problems that I could not download the dataset successfully. Thus, we need to add this command.
options(repos = c(CRAN = "https://cloud.r-project.org"))

# To make sure all outputs are english, rather than Mandarin.
Sys.setenv(LANG = "en")
```

```{r}
# Download the package

install.packages("dplyr")

install.packages("ggplot2")

library("dplyr")

library("ggplot2")

```

## 1.Data Preparation

```{r}
# Import Data (FoodAtlas.csv)
food_data <- read.csv("FoodAtlas.csv")

# show only the first 24 
# head(food_data, 24)

# show only last 12 variables
# tail(food_data,12)

# For keep only the first 24 and last 12 variables.

selected_variables <- c(colnames(food_data)[1:24], colnames(food_data)[(length(colnames(food_data)) - 11):length(colnames(food_data))])

# reduced dataset is named as reduces_data
reduced_data <- food_data[, selected_variables]

# show the data      
head(reduced_data)            
```

## 2.Examine this reduced dataset and decribe their types.

```{r}
## Food_Data is a Data Frame.
# What's the data type
class(food_data)

## And the object's type is list
typeof(food_data)

# Check for the reduced data
str(reduced_data)

# Decribe the reduced data
summary(reduced_data)
```

## 3. Make a new data object for the state of Kansas with specific features.

```{r}

## Make a new data object named "ks" for the state of Kansas.
ks_data <- subset(food_data, State == "Kansas")

head(ks_data)

# Keep 6 specific features
ks_data <- select(ks_data, County, POP2010, LAhalfand10, MedianFamilyIncome, TractSNAP, TractKids )

head(ks_data)

# super version
# ks_data <- food_data[food_data$State == "Kansas", c("County", "POP2010", "LAhalfand10", "MedianFamilyIncome", "TractSNAP", "TractKids")]

```

## 4. Create a plot to illustrate the differences in across counties in Kansas.

```{r}

# Using ggplot2 to draw a boxplot.
ggplot(ks_data, aes(x = County, y = MedianFamilyIncome)) +
  geom_boxplot() +
  coord_flip() +
  labs(x = "County", y = "Median Family Income", title = "Income Distribution Across Counties in Kansas")

# save the image to check detail
ggsave("4.income_distribution_plot.png", plot = last_plot(), width = 20, height = 16, dpi = 300)
```

## From the boxplot image, we can find Johnson own the greatest variation in income.

## Are there significant differences across counties?

Answer：1.Variation in Median Income: The middle of boxplot means the
median income, and different income exists in counties. Some of them are
high, others are low.

Answer：2.Variation in Interquartile Range: The boxes dedicate the
interquartile range, as well as the middle 50% of incomes. There are
huge differences in the interquartile range across difference counties,
meaning differences in income distribution.

```{r}
# Calculate the range of Median Family Income
county_variation <- aggregate(MedianFamilyIncome ~ County, data = ks_data, FUN = function(x) var(x, na.rm = TRUE))

# Find the the greatest variation in income to double check the viewpoint.
county_max_variation <- county_variation[which.max(county_variation$MedianFamilyIncome), ]

## Which counties have the greatest variation in income?
print(paste("Answer：County with the greatest variation in income:", county_max_variation$County))

```

## 5. How compares across locations with low (or not) food accessibility ().

```{r}
# Compare using both basic descriptive statistics and graphical representation
summary(ks_data$MedianFamilyIncome[ks_data$LAhalfand10 == 1])  
summary(ks_data$MedianFamilyIncome[ks_data$LAhalfand10 == 0])  

# Using ggplot2 to compare
ggplot(ks_data, aes(x = MedianFamilyIncome, fill = factor(LAhalfand10))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
  labs(x = "Median Family Income", y = "Count", fill = "Low Food Accessibility") +
  ggtitle("Distribution of Median Family Income by Food Accessibility") +
  scale_fill_manual(values = c("blue", "red"))

```

## Summary Statistics:

-   For areas with low food accessibility:

1.  Minimum Median Family Income: 0
2.  Maximum Median Family Income: 225,739
3.  Median Median Family Income: \$59,107
4.  Mean Median Family Income: \$63,992

-   For areas with better food

1.  Minimum Median Family Income: 0
2.  Maximum Median Family Income: 117,703
3.  Median Median Family Income: \$63,125
4.  Mean Median Family Income: \$62,653

## Summary histogram:

-   Both of the classes (LAhalfand10 == 0 or 1), there is no data
    located at approximately 200000 in Median Family Income. And most of
    data concentrate within 25000-125000.

-   The histogram for areas with low food accessibility (LAhalfand10 ==

    1)  shows a slightly right-skewed distribution, with a peak around
        the median value and a long tail to the right. The histogram for
        areas with better food accessibility (LAhalfand10 == 0) also
        exhibits a slightly right-skewed distribution, with a peak
        around the median value, but the tail to the right is less
        pronounced compared to areas with low food accessibility.

## 6. Using other variable to compare between counties with TractKids

```{r}
# Compare using both basic descriptive statistics and graphical representation
summary_median_income_low_access <- summary(ks_data$MedianFamilyIncome[ks_data$LAhalfand10 == 1])
summary_median_income_regular_access <- summary(ks_data$MedianFamilyIncome[ks_data$LAhalfand10 == 0])

# Using ggplot2 to compare TractKids
ggplot(ks_data, aes(x = TractKids, fill = factor(LAhalfand10))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
  labs(x = "TractKids", y = "Count", fill = "Food Accessibility") +
  ggtitle("Distribution of TractKids by Food Accessibility") +
  scale_fill_manual(values = c("blue", "red"))


summary_median_income_low_access
summary_median_income_regular_access
```

## Summary Statistics:

-   For areas with low food accessibility (LAhalfand10 == 1):

1.  Minimum Median Family Income: \$0
2.  Median Median Family Income: \$59,107
3.  Mean Median Family Income: \$63,992
4.  Maximum Median Family Income: \$225,739

-   For areas with regular food accessibility (LAhalfand10 == 0):

1.  Minimum Median Family Income: \$0
2.  Median Median Family Income: \$63,125
3.  Mean Median Family Income: \$62,653
4.  Maximum Median Family Income: \$117,703

## Summary histogram:

-   Apparently, the count of food accessibility == 1 is higher than food
    accessibility == 0. Second, we can see that for all countries in
    Kansas, the median family income mostly concentrates in lower
    intervals. This indicates that the majority of counties have
    relatively lower median family incomes.

## 7. Explanation of Kansas dataset relate to food accessibility and food assistance (SNAP).

```{r}
# Compare using both basic descriptive statistics and graphical representation
summary(ks_data$TractSNAP[ks_data$LAhalfand10 == 1])  
summary(ks_data$TractSNAP[ks_data$LAhalfand10 == 0])  

# Using ggplot2 to compare
ggplot(ks_data, aes(x = TractSNAP, fill = factor(LAhalfand10))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
  labs(x = "SNAP Recipients", y = "Count", fill = "Low Food Accessibility") +
  ggtitle("Distribution of SNAP Recipients by Food Accessibility") +
  scale_fill_manual(values = c("blue", "red"))
```

## Summary Statistics:

-   For areas with low food accessibility (LAhalfand10 == 1):

1.  Minimum number of SNAP recipients: 0
2.  Maximum number of SNAP recipients: 686
3.  Median number of SNAP recipients: 113
4.  Mean number of SNAP recipients: 146.4

-   For areas with regular food accessibility (LAhalfand10 == 0):

1.  Minimum number of SNAP recipients: 0
2.  Maximum number of SNAP recipients: 509
3.  Median number of SNAP recipients: 93
4.  Mean number of SNAP recipients: 109.3

## Summary histogram:

-   Obviouly, the count of LAhalfand10 == 1 is higher then LAhalfand10
    == 0 significantly. Secondly, most distribution concentrate in
    0-400.

## 8. Use reduced_data created in step 1 and create a state-by-state summary of each variable.

```{r}
# Grouping data by State
state_summary <- reduced_data %>%
  group_by(State) %>%
  summarize(
    Mean_Population = mean(POP2010),
    Mean_Occupied_Housing_Units = mean(OHU2010),
    Mean_Poverty_Rate = mean(PovertyRate),
    Median_Family_Income = median(MedianFamilyIncome),
    Mean_LowIncome_Tracts = mean(LowIncomeTracts),
    Mean_Low_Access_1And10 = mean(LA1and10),
    Mean_Low_Access_halfAnd10 = mean(LAhalfand10),
    Mean_Low_Access_1And20 = mean(LA1and20),
    Mean_Low_Access_Tracts_Half = mean(LATracts_half),
    Mean_Low_Access_Tracts_1 = mean(LATracts1),
    Mean_Low_Access_Tracts_10 = mean(LATracts10),
    Mean_Low_Access_Tracts_20 = mean(LATracts20),
    Mean_Households_With_No_Vehicle = mean(HUNVFlag),
    Mean_Snap_Participation = mean(TractSNAP)
  )

# Identify states with the highest and lowest food accessibility problems
highest_problems <- state_summary %>%
  arrange(desc(Mean_LowIncome_Tracts)) %>%
  head(5)

lowest_problems <- state_summary %>%
  arrange(Mean_LowIncome_Tracts) %>%
  head(5)

# Print results
cat("States with the most problems with food accessibility:\n")
print(highest_problems)

cat("\nStates with the lowest problems with food accessibility:\n")
print(lowest_problems)

```


## 9. Examine the trends between 3 variables of your choice and food accessibility across states in the US and describe your findings.

Following plot would focus on PovertyRate, MedianFamilyIncome, and HUNVFlag.

```{r}
# Scatterplot matrix to examine relationships
scatterplot_matrix <- ggplot(state_summary, aes(x = Mean_Poverty_Rate, y = Mean_LowIncome_Tracts)) +
  geom_point(aes(color = State)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship between Poverty Rate and Food Accessibility",
       x = "Poverty Rate (%)",
       y = "Mean Percentage of Low-Income Tracts (%)") +
  theme_minimal()

scatterplot_matrix

# Scatterplot matrix for Median Family Income vs. Food Accessibility
scatterplot_matrix_income <- ggplot(state_summary, aes(x = Median_Family_Income, y = Mean_LowIncome_Tracts)) +
  geom_point(aes(color = State)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship between Median Family Income and Food Accessibility",
       x = "Median Family Income ($)",
       y = "Mean Percentage of Low-Income Tracts (%)") +
  theme_minimal()

scatterplot_matrix_income

# Scatterplot matrix for Percentage of households with no vehicle vs. Food Accessibility
scatterplot_matrix_vehicle <- ggplot(state_summary, aes(x = Mean_Households_With_No_Vehicle, y = Mean_LowIncome_Tracts)) +
  geom_point(aes(color = State)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship between Percentage of Households with No Vehicle and Food Accessibility",
       x = "Percentage of Households with No Vehicle (%)",
       y = "Mean Percentage of Low-Income Tracts (%)") +
  theme_minimal()

scatterplot_matrix_vehicle

```

1.  The first scatterplot matrix shows the relationship between poverty rate and food accessibility. It seems there's a positive correlation, indicating that states with higher poverty rates tend to have a higher percentage of low-income tracts, suggesting poorer food accessibility.


2.  The second scatterplot matrix explores the relationship between median family income and food accessibility. There appears to be a negative correlation, implying that states with higher median family incomes tend to have a lower percentage of low-income tracts and potentially better food accessibility. But! This regression is not strong relative as first scatterplot. There are more states be away from the regression line.


3.  The third scatterplot matrix investigates the relationship between the percentage of households with no vehicle and food accessibility. It shows a positive correlation, suggesting that states with a higher percentage of households with no vehicle may experience challenges with food accessibility, potentially due to transportation limitations. However, the correlation is weakest.

## In this comparative analysis involving three variables—PovertyRate, MedianFamilyIncome, and HUNVFlag—it is evident that the regression line generated by PovertyRate exhibits the tightest clustering of data points. Following this, MedianFamilyIncome shows a moderate level of dispersion, while HUNVFlag displays the least concentration of data points, indicating the weakest correlation.

# Hence, among the three variables, Mean PovertyRate emerges as the most influential factor affecting the Percentage of Low-Income Tracts.
