---
title: 'Lab #3'
author: "WEI-XUAN HU"
date: "2024-02-05"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## 0. Download packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# There are some problems that I could not download the dataset successfully. Thus, we need to add this command.
options(repos = c(CRAN = "https://cloud.r-project.org"))

# To make sure all outputs are english, rather than Mandarin.
Sys.setenv(LANG = "en")
```

```{r}
# Download the package

install.packages("dplyr")

install.packages("ggplot2")

library("dplyr")

library("ggplot2")

```

## 1.Data Preparation

```{r}
# Import Data (FoodAtlas.csv)
food_data <- read.csv("FoodAtlas.csv")

# show only the first 24 
# head(food_data, 24)

# show only last 12 variables
# tail(food_data,12)

# For keep only the first 24 and last 12 variables.

selected_variables <- c(colnames(food_data)[1:24], colnames(food_data)[(length(colnames(food_data)) - 11):length(colnames(food_data))])

# reduced dataset is named as reduces_data
reduced_data <- food_data[, selected_variables]
                  
```

## 2.Examine this reduced dataset and decribe their types.

```{r}
## Food_Data is a Data Frame.
# What's the data type
class(food_data)

## And the object's type is list
typeof(food_data)

# Check for the reduced data
str(reduced_data)

# Decribe the reduced data
summary(reduced_data)
```

## 3. Make a new data object for the state of Kansas with specific features.

```{r}

## Make a new data object named "ks" for the state of Kansas.
ks_data <- subset(food_data, State == "Kansas")

head(ks_data)

# Keep 6 specific features
ks_data <- select(ks_data, County, POP2010, LAhalfand10, MedianFamilyIncome, TractSNAP, TractKids )

head(ks_data)

# super version
# ks_data <- food_data[food_data$State == "Kansas", c("County", "POP2010", "LAhalfand10", "MedianFamilyIncome", "TractSNAP", "TractKids")]

```

## 4. Create a plot to illustrate the differences in across counties in Kansas.

```{r}

# Using ggplot2 to draw a boxplot.
ggplot(ks_data, aes(x = County, y = MedianFamilyIncome)) +
  geom_boxplot() +
  coord_flip() +
  labs(x = "County", y = "Median Family Income", title = "Income Distribution Across Counties in Kansas")

# save the image to check detail
ggsave("4.income_distribution_plot.png", plot = last_plot(), width = 20, height = 16, dpi = 300)
```

## From the boxplot image, we can find Johnson own the greatest variation in income.

## Are there significant differences across counties?

Answer：1.Variation in Median Income: The middle of boxplot means the
median income, and different income exists in counties. Some of them are
high, others are low.

Answer：2.Variation in Interquartile Range: The boxes dedicate the
interquartile range, as well as the middle 50% of incomes. There are
huge differences in the interquartile range across difference counties,
meaning differences in income distribution.

```{r}
# Calculate the range of Median Family Income
county_variation <- aggregate(MedianFamilyIncome ~ County, data = ks_data, FUN = function(x) var(x, na.rm = TRUE))

# Find the the greatest variation in income to double check the viewpoint.
county_max_variation <- county_variation[which.max(county_variation$MedianFamilyIncome), ]

## Which counties have the greatest variation in income?
print(paste("Answer：County with the greatest variation in income:", county_max_variation$County))

```

## 5. How compares across locations with low (or not) food accessibility ().

```{r}
# Compare using both basic descriptive statistics and graphical representation
summary(ks_data$MedianFamilyIncome[ks_data$LAhalfand10 == 1])  
summary(ks_data$MedianFamilyIncome[ks_data$LAhalfand10 == 0])  

# Using ggplot2 to compare
ggplot(ks_data, aes(x = MedianFamilyIncome, fill = factor(LAhalfand10))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
  labs(x = "Median Family Income", y = "Count", fill = "Low Food Accessibility") +
  ggtitle("Distribution of Median Family Income by Food Accessibility") +
  scale_fill_manual(values = c("blue", "red"))

```

## Summary Statistics:

-   For areas with low food accessibility:

1.  Minimum Median Family Income: 0
2.  Maximum Median Family Income: 225,739
3.  Median Median Family Income: $59,107
4.  Mean Median Family Income: $63,992

-   For areas with better food

1.  Minimum Median Family Income: 0
2.  Maximum Median Family Income: 117,703
3.  Median Median Family Income: $63,125
4.  Mean Median Family Income: $62,653

## Summary histogram:

-   Both of the classes (LAhalfand10 == 0 or 1), there is no data located at approximately 200000 in Median Family Income.
    And most of data concentrate within 25000-125000.

-   The histogram for areas with low food accessibility (LAhalfand10 ==
    1) shows a slightly right-skewed distribution, with a peak around
    the median value and a long tail to the right. The histogram for
    areas with better food accessibility (LAhalfand10 == 0) also
    exhibits a slightly right-skewed distribution, with a peak around
    the median value, but the tail to the right is less pronounced
    compared to areas with low food accessibility.

## 6. Using other variable to compare between counties with TractKids

```{r}
# Compare using both basic descriptive statistics and graphical representation
summary_median_income_low_access <- summary(ks_data$MedianFamilyIncome[ks_data$LAhalfand10 == 1])
summary_median_income_regular_access <- summary(ks_data$MedianFamilyIncome[ks_data$LAhalfand10 == 0])

# Using ggplot2 to compare TractKids
ggplot(ks_data, aes(x = TractKids, fill = factor(LAhalfand10))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
  labs(x = "TractKids", y = "Count", fill = "Food Accessibility") +
  ggtitle("Distribution of TractKids by Food Accessibility") +
  scale_fill_manual(values = c("blue", "red"))


summary_median_income_low_access
summary_median_income_regular_access
```

## Summary Statistics:

-   For areas with low food accessibility (LAhalfand10 == 1):

1.  Minimum Median Family Income: $0
2.  Median Median Family Income: $59,107
3.  Mean Median Family Income: $63,992
4.  Maximum Median Family Income: $225,739

-   For areas with regular food accessibility (LAhalfand10 == 0):

1.  Minimum Median Family Income: $0
2.  Median Median Family Income: $63,125
3.  Mean Median Family Income: $62,653
4.  Maximum Median Family Income: $117,703

## Summary histogram:

-   Apparently, the count of food accessibility == 1 is higher than food accessibility == 0. Second, we can  see that for all countries in Kansas, the median family income mostly concentrates in lower intervals. This indicates that the majority of counties have relatively lower median family incomes.

## 7. Explanation of Kansas dataset relate to food accessibility and food assistance (SNAP).


```{r}
# Compare using both basic descriptive statistics and graphical representation
summary(ks_data$TractSNAP[ks_data$LAhalfand10 == 1])  
summary(ks_data$TractSNAP[ks_data$LAhalfand10 == 0])  

# Using ggplot2 to compare
ggplot(ks_data, aes(x = TractSNAP, fill = factor(LAhalfand10))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
  labs(x = "SNAP Recipients", y = "Count", fill = "Low Food Accessibility") +
  ggtitle("Distribution of SNAP Recipients by Food Accessibility") +
  scale_fill_manual(values = c("blue", "red"))
```

## Summary Statistics:

-   For areas with low food accessibility (LAhalfand10 == 1):

1.  Minimum number of SNAP recipients: 0
2.  Maximum number of SNAP recipients: 686
3.  Median number of SNAP recipients: 113
4.  Mean number of SNAP recipients: 146.4

-   For areas with regular food accessibility (LAhalfand10 == 0):

1.  Minimum number of SNAP recipients: 0
2.  Maximum number of SNAP recipients: 509
3.  Median number of SNAP recipients: 93
4.  Mean number of SNAP recipients: 109.3

## Summary histogram:

-   Obviouly, the count of LAhalfand10 == 1 is higher then LAhalfand10 == 0 significantly. Secondly, most distribution         concentrate in 0-400.


## 8. Use reduced_data created in step 1 and create a state-by-state summary of each variable.

state_summary <- reduced_data %>%
  group_by(State) %>%
  summarise_all(list(mean = mean, median = median, sd = sd, min = min, max = max))
  
# Display the state-by-state summary
print(state_summary)





















