---
title: 'Lab #2'
author: "WEI-XUAN HU"
date: "2024-01-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# There are some problems that I could not download the dataset successfully. Thus, we need to add this ccommand.
options(repos = c(CRAN = "https://cloud.r-project.org"))

# To make sure all outputs are english, rather than Mandarin.
Sys.setenv(LANG = "en")
```

## 1.Data Preparation

```{r}
# Download the package and dataset
install.packages("gapminder")

install.packages("dplyr")

install.packages("ggplot2")

library("gapminder")

library("dplyr")

library("ggplot2")

data(gapminder)
 
head(gapminder, 15)

# The data order countries in reverse alphabetical order 
gm <- arrange(gapminder, desc(country))

# Display the first 15
head(gm, 15)

```

## Ordering problems in a different dataset

1.  Non-Alphanumeric Data:

If the data are not alphanumeric characters like special symbols or other language, the reverse order would not work.

2.  Missing Values:

If there are missing values in the column used for ordering, they might affect the results.

## 2.Calculate and display the mean, median, and standard deviation

```{r}
# output is the mean for the dataset.
mean_lifeExp <- mean(gm$lifeExp)

mean_pop <- mean(gm$pop)

mean_gdpPercap <- mean(gm$gdpPercap)

cat("mean of Life expectancy:", mean_lifeExp, "\n")

cat("mean of Life Population:", mean_pop, "\n")

cat("mean of Life GDPperCapita:", mean_gdpPercap, "\n")


# output is the median for the dataset.
median_lifeExp <- median(gm$lifeExp)

median_pop <- median(gm$pop)

median_gdpPercap <- median(gm$gdpPercap)


cat("median of Life expectancy:", median_lifeExp, "\n")

cat("median of Life Population:", median_pop, "\n")

cat("median of Life GDPperCapita:", median_gdpPercap, "\n")


# output is the standard deviation of the dataset.
sd_lifeExp <- sd(gm$lifeExp)

sd_pop <- sd(gm$pop)

sd_gdpPercap <- sd(gm$gdpPercap)


cat("standard deviation  of Life expectancy:", sd_lifeExp, "\n")

cat("standard deviation  of Life Population:", sd_pop, "\n")

cat("standard deviation  of Life GDPperCapita:", sd_gdpPercap, "\n")

```

## 3.Calculate and display in a single dataframe the mean values for Life expectancy, Population, and GDPperCapita for each continent.

```{r}
# Calculate the mean values for each continent
continent_means <- gapminder %>%
  
  group_by(continent) %>%
  
  summarise(
    
    mean_lifeExp, mean_pop, mean_gdpPercap
    
  )

# Print the resulting dataframe
print(continent_means)

```

## 4.Create a scatter plot of the global average Life expectancy over time and display the first 5 records for the dataset used to build this plot. Customize the display of the plot to make it less blah.

```{r}
# Calculate global average life expectancy over time
global_avg_life_expectancy <- gapminder %>%
  
  group_by(year) %>%
  
  summarise(Avg_LifeExpectancy = mean(lifeExp, na.rm = TRUE))


# Create scatter plot

scatter_plot <- ggplot(global_avg_life_expectancy, aes(x = year, y = Avg_LifeExpectancy)) +
  
  geom_point(size = 3, color = "steelblue") +
  
  geom_smooth(method = "loess", se = FALSE, color = "darkred", linewidth = 1) +
  
  labs(
    
    title = "Global Average Life Expectancy Over Time",
    
    x = "Year",
    
    y = "Average Life Expectancy"
    
  ) +
  
  theme_minimal()

# Display the first 5 records of the dataset

head(global_avg_life_expectancy, 5)

# Display the scatter plot
print(scatter_plot)

```

## 5.Pull in a dataset of your choice from the datasets package that has a dataframe structure (not sure what’s available, try ?datasets). Then examine the structure of the data including: class types, header information, and dimensions.

```{r}
# Load the iris dataset
data(iris)

# Display the structure of the dataset
str(iris)

# Display the first few rows of the dataset
head(iris)

# Display the dimensions of the dataset
dim(iris)

```

## 6.Calculate general summary statistics for the dataset.

```{r}
# Calculate general summary statistics for the "iris" dataset
summary(iris)

```

## 7.Calculate summary statistics for groups/subsets of the dataset (these groupings could be based on time period, species, location, etc…).

```{r}
# Calculate summary statistics for groups/subsets based on species
summary_by_species <- iris %>%
  
  group_by(Species) %>%
  
  summarise(
    
    Mean_SepalLength = mean(Sepal.Length),
    
    Mean_SepalWidth = mean(Sepal.Width),
    
    Mean_PetalLength = mean(Petal.Length),
    
    Mean_PetalWidth = mean(Petal.Width),
    
    Median_SepalLength = median(Sepal.Length),
    
    Median_SepalWidth = median(Sepal.Width),
    
    Median_PetalLength = median(Petal.Length),
    
    Median_PetalWidth = median(Petal.Width),
    
    Min_SepalLength = min(Sepal.Length),
    
    Min_SepalWidth = min(Sepal.Width),
    
    Min_PetalLength = min(Petal.Length),
    
    Min_PetalWidth = min(Petal.Width),
    
    Max_SepalLength = max(Sepal.Length),
    
    Max_SepalWidth = max(Sepal.Width),
    
    Max_PetalLength = max(Petal.Length),
    
    Max_PetalWidth = max(Petal.Width)
    
  )

# Print the summary statistics by species
print(summary_by_species, width = Inf)

```

## 8.Create 2 plots of your choice, one showing grouped data and the other showing ungrouped data. Give each plot a title and descriptive x and y axis labels. Customize at least one graphic option (color, point type, line width, etc…) on these plots.

```{r}
# Plot 1: Grouped Data (Scatter plot)
plot_grouped <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  
  geom_point(size = 3, alpha = 0.8) +
  
  labs(
    
    title = "Scatter Plot of Sepal Dimensions (Grouped by Species)",
    
    x = "Sepal Length",
    
    y = "Sepal Width"
    
  ) +
  
  theme_minimal()


# Plot 2: Ungrouped Data (Box plot)
plot_ungrouped <- ggplot(iris, aes(x = Species, y = Petal.Length, fill = Species)) +
  
  geom_boxplot(width = 0.5, alpha = 0.8) +
  
  labs(
    
    title = "Box Plot of Petal Length (Ungrouped)",
    
    x = "Species",
    
    y = "Petal Length"
    
  ) +
  
  theme_minimal()


# Display the plots
print(plot_grouped)

print(plot_ungrouped)
```
