---
title: 'Lab #5'
author: "WEI-XUAN HU"
date: "2024-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# There are some problems that I could not download the dataset successfully. Thus, we need to add this command.
options(repos = c(CRAN = "https://cloud.r-project.org"))

# To make sure all outputs are english, rather than Mandarin.
Sys.setenv(LANG = "en")
```

#  {.tabset}

## 0.
### Install packages required.

```{r}
install.packages("tidyverse")
install.packages("sf")
install.packages("spData")
install.packages("ggplot2")


library(tidyverse)
library(sf)
library(spData)
library(ggplot2)
```

## 1.
### Load the dataset and combine them

```{r}

data1 <- world
data2 <- worldbank_df

str(data1)
head(data1)

str(data2)
head(data2)

combined_data <- merge(data1, data2)

str(combined_data)

```

We can see that data1 is in sf format and data2 is in tibble format. When these two are merged, the resulting format will be a data frame. This is because data1 contains a column 'geom' in sf format, while data2 does not have any 'geom' attribute, and none of the attributes in data2 belong to sf format. Therefore, the combined_data will not be in sf format.


## 2.
### Make some basic (non-spatial) plots to explore differences in urban population and literacy across countries.

```{r}
ggplot(combined_data, aes(x = urban_pop, y = literacy)) +
  geom_point() +
  labs(x = "Urban Population", y = "Literacy Rate", title = "Urban Population vs. Literacy Rate")
```

## 3.
### Now make some maps to explore urban population and literacy across countries.

```{r}
# Plot urban population on a map
plot(combined_data["urban_pop"], main = "Urban Population Across Countries")

# Plot literacy rate on a map
plot(combined_data["literacy"], main = "Literacy Rate Across Countries")

```

## 4.
### Make some basic (non-spatial) plots to explore differences in urban population and literacy across continents.

```{r}
ggplot(combined_data, aes(x = urban_pop, y = literacy, color = region_un)) +
  geom_point() +
  labs(x = "Urban Population", y = "Literacy Rate", title = "Urban Population vs. Literacy Rate Across Continents")
```

## 5.
### Now make some maps to explore differences in urban population and literacy across continents. Change to a different, projected coordinate system for these maps.

```{r}
# Project the data to a different coordinate system：Robinson projection
combined_data_rob_proj <- st_transform(combined_data, "+proj=robin")

# Plot urban population on a map
ggplot() +
  geom_sf(data = combined_data_rob_proj, aes(fill = urban_pop)) +
  scale_fill_viridis_c() +
  labs(title = "Urban Population Across Continents (Projected)") +
  theme_minimal()

# Plot literacy rate on a map
ggplot() +
  geom_sf(data = combined_data_rob_proj, aes(fill = literacy)) +
  scale_fill_viridis_c() +
  labs(title = "Literacy Rate Across Continents (Projected)") +
  theme_minimal()

```